Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-09 edubots_surveys.Rnw
Review methods - par 1,2

\begin{abstract}
  Messaging platforms are applications, generally mediated by an app, desktop
  program or the web, mainly used for synchronous communication among users. As
  such, they have been widely adopted officially by higher education
  establishments, after little or no study of their impact and perception by the
  teachers. We think that the introduction of these new tools and the
  opportunities and challenges they have needs to be studied carefully in order
  to adopt the model, as well as the tool, that is the most adequate for all
  parties involved. We already studied the perception of these tools by
  students, in this paper we examine the teachers’ experiences and perceptions
  through a survey that we validated with peers, and what they think these tools
  should make or serve so that it enhances students learning and helps them
  achieve their learning objectives. The survey has been distributed among
  tertiary education teachers, both in universitary and other kind of tertiary
  establishments, based in Spain (mainly) and Spanish-speaking countries. We
  have focused on collecting teachers' preferences and opinions on the
  introduction of messaging platforms in their day-to-day work, as well as other
  services attached to them, such as chatbots. What we intend with this survey
  is to understand their needs and to gather information about the various
  educational use cases where these tools could be valuable. In addition, an
  analysis of how and when teachers’ opinions towards the use of these tools
  varies across gender, experience, and their discipline of specialization is
  presented. The key findings of this study highlight the factors that can
  contribute to the advancement of the adoption of messaging platforms and
  chatbots in higher education institutions to achieve the desired learning
  outcomes.
\end{abstract}

@@

However, the success of this adoption largely depends on the willingness and
ability of teachers to embrace it, as they are the primary agents in this new
educational approach. As such, gathering and assessing feedback from these users
will be crucial in realizing the potential benefits of the technology.

@@

Given the potential benefits that chatbots could bring to the classroom, the
focus of the % EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
unnamed European project, the main sponsor of this study, is to explore best
practices and innovative use of chatbots, and to create a learning community of
educators in higher education institutions.  Researchers and teachers
participating in this project have developed several studies. Most of these
works have been focused on how messaging applications and/or chatbots are
normally applied to deliver personalized learning in classrooms that occurs
anytime anywhere, and promote collaborative learning experiences or group
discussions \cite{panah2020study}, and boost students' sense of belonging to
their institutions \cite{abbas2021onlinechat}.  The majority of these studies
were conducted focusing on the teachers' perspectives, so this paper aims to
provide more conclusions regarding their opinions and needs, as well as the
existing challenges or opportunities that arise when adopting these technologies
in classrooms. This will lead to clarify how they can positively contribute to
enhancing the learning process in higher education institutions.

In this study, data collection involved two phases:
 \begin{itemize}
 \item The first phase aimed to collect feedback from students about their
   students’ preferences of tools/applications for chatting and messaging, how
   they use them in educational context, who they like to be with in the class
   messaging groups, and their expectations of chatbots in assisting them during
   their learning process. Therefore, two surveys for bachelor and master degree
   students at the University of Granada (Spain) were designed and answers from
   more than 250 students were collected. The key results from the student
   surveys indicate a preference for using familiar messaging applications, such
   as Telegram or WhatsApp. Students also expressed interest in using chatbots
   that can assist them with organizing their course schedules, help them access
   their assignment grades and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed; this
  will be detailed in Section \ref{sec:meth}. The responses' analysis led to
the preparation of a second survey for teachers and 200 responses were
collected, again following the methodology that will be
  explained later. The questions of the second survey focused on asking about
  teachers' needs, opinions and preferences of the development of future
  technology-enhanced tools and its potential impact on educational institutions
  policies. In order to assess these, this paper layed out the following
  research questions:

@@

We will try to respond to these research questions through the analysis of the
answers provided by professors to the two aforementioned surveys.  Eventually,
the main research outcome of this survey should be a series of recommendations
to make a successful deployment of chatbot technologies (and, in some cases,
general instant messaging applications) in higher education; from the response
to RQ1 and RQ4 we will try to recommend specific technologies or applications to
be deployed in the classroom; the response to RQ2 and RQ3 will help us recommend
chatbots features or specific platforms; and finally, from RQ3 we will also try
to find best practices in the adoption of messaging platforms and their matching
chatbots.

@@

The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education institutions facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries, the most popular MIM platform. The reason is that it is the tool preferably
used by educators to provide assignments' feedback to students, support course
discussions, and post learning resources in a semi-formal learning context
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education, as well as applications based on them such as chatbots, is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell in
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
using chatbots to collect course feedback from students in higher
education improved the quality of responses given by students in their assignments, and boosted engagement levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

Several studies to evaluate chatbots in higher education have been conducted. For
instance, a comprehensive study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, concluding that most of these chatbots were very basic and that programmers require of more support to develop and 
offer tools for improving coaching methods and students learning. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which respond to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada deployed
Elvira, a chatbot embedded in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the speech-mouth and face gestures synchronization rather than making updating
answered questions simpler or more interactive for the administrative staff, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

Contrarily to the mentioned chatbots that are focused on administrative tasks, teaching-oriented chatbots are more sophisticated, as they
set personalized learning outcomes and monitor learning progress. For instance,
\cite{fernoagua2018intelligent} reported on ``eduAssistant'', a virtual teaching
assistant chatbot developed on the Telegram messaging platform. In this study,
the Telegram platform was chosen because it is easy to use, students are familiar
with its features, and it enables them to exchange messages in different formats
(text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram
can operate on all devices and operating systems. The ``eduAssistant'' chatbot
acts as an automatic agent in the teacher-content-student relationship, providing real-time
feedback loops and a personalized learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose questions to
their students and the chatbot assists those who need further help by giving
them more hints and reporting it to the educator's dashboard
\cite{fernoagua2018intelligent}. This could help educators identify
students who need more attention and provide them with more educational resources
relevant to their academic achievement. %In addition, a recent study
%has analysed how the use of chatbots positively affected the learning outcomes
%of students in a Chinese class \cite{doi:10.1177/0735633120929622}.

Despite its (arguably) successful implementation in different higher education
institutions, particularly on the fringes of educational activity rather than actual
student-teacher interaction, its implementation or deployment is not trivial.
Some authors, for instance \cite{sjostrom2018designing}, have proposed a
conceptual architecture for the adoption of teaching-oriented chatbots in higher
education.  This conceptual architecture is based on a systematic literature
review of previous studies examining the design of chatbots in higher
education, as well as making a content analysis of student emails and discussion
forum posts from four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasized the importance
of developing chatbots in platforms that students and educators are familiar
with and can easily access (i.e. Facebook Messenger), which was confirmed by
\cite{hobert2019you} and \cite{fernoagua2018intelligent}. In addition, \cite{sjostrom2018designing} argued that 
conceptualizing learners' questions could help designers integrate the
appropriate types of questions that chatbots should support for different
courses. Other authors, such as \cite{coronado2018cognitive}, have proposed agents that
record learning materials to be provided on demand to students, while \cite{crockett2017predicting} 
reported on tutoring systems that can
perform initial assessments of students' understanding and provide learning
materials to advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many
studies have focused on the evaluation of technology acceptance and usability
\cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain
where, according to \cite{hobert2019you}, specific pedagogical factors
such as learning success and increased motivation are more important.  Therefore,
to develop effective chatbots for higher education, the needs of all stakeholders
(i.e. educators, students, institutions, etc) should be carefully
collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}. These needs include, but are not limited to, student learning success due to higher motivation, but these are a posteriori effects that cannot be assessed in advance. Both authors focus on what teachers need in terms of the latter; this paper will focus on a wider perspective, trying to determine what they are looking for in terms of general messaging technology. The classical literature \cite{tam91} already proposes that any adoption of technology must be tuned to the user needs and experience. In this paper, we will try to find out what those are in order to propose a successful model of adoption of chatbot technology.

In line with this research, our previous work \cite{MoraChatbots2021:anon}
aimed at analyzing the expectations of students in this regard. Another paper
have also analysed how the use of chatbots affected the learning outcomes of
students in a Chinese class \cite{doi:10.1177/0735633120929622}. This paper
considered conversational chatbots in a one-to-one setting, finding that learning
truly benefited from it. This is, once again, an evidence of the benefit of
chatbots in certain settings; however, there are some prior experiences, as well as needs, that may
prevent the successful acceptance of the technology, and these are what we are
trying to find out in this paper, along with what kind of features would improve
its acceptance.

@@

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Kind of interactions} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 166 & 59.1 & 115& 40.9 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers of the School/Faculty\end{tabular} & 19 & 6.7 & 266 & 94.3 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 58 & 20.5 & 224 & 79.4 \\
\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 81 & 28.7 & 201 & 71.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Kind of interactions preferred}
\label{tab:kindinteractions}
\end{table}

@@

@@

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

This inclusion of university-provided messaging applications in
the survey might explain why there are so few teachers, just 6.7\%, that answered they did not
use any messaging application. In general, universities provide official channels
to send students academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
teaching staff might not use it, but use some extraofficial or opt-in channels such as the others included in the survey; in the end, between these official and extra-official channels, most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions arising from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots, and just a few subjects selected
the option "Others". This may suggest that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive. If the responders were more acquainted with the possibilities of this technology, this response would possibly have shown up many more times. Another way to look at it is that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burnt out to think about innovation.

We can speculate that teaching staff's opinions of chatbots are influenced by negative
experiences with chatbots used in customer support, which is what most people are likely to have
experienced. Further study may be needed to confirm this, but it is important to 
keep in mind when introducing chatbot technologies in the classroom.
To potentially change this existing perception, it will be necessary to make chatbots as simple and satisfying to use as possible.

>>>>

\begin{abstract} % abstract
Messaging platforms are applications, generally mediated by an app, desktop program or the web, mainly used for synchronous communication among users. As such, they have been widely adopted officially by higher education establishments, after little or no study of their impact and perception by the teachers. We think that the introduction of these new tools and the opportunities and challenges they have needs to be studied carefully in order to adopt the model, as well as the tool, that is the most adequate for all parties involved. We already studied the perception of these tools by students, in this paper we examine the teachers’
experiences and perceptions through a survey that we validated with peers, and what they think these tools should make or serve so that it enhances students learning and helps them achieve their learning objectives. The survey has been distributed among tertiary education teachers, both in universitary and other kind of tertiary establishments, based in
Spain (mainly) and Spanish-speaking countries. We have focused on collecting
teachers' preferences and opinions on the introduction of messaging platforms in their day-to-day work, as well as other services attached to them, such as  chatbots. What we intend with this survey is to understand their needs and to gather
information about the various educational use cases where these tools could be
valuable. In addition, an analysis of how and when teachers’ opinions towards
the use of these tools varies across gender, experience, and their discipline
of specialization is presented. The key findings of this study highlight the
factors that can contribute to the advancement of the adoption of messaging
platforms and chatbots in higher education institutions to achieve the desired
learning outcomes.
\end{abstract}

@@

Anyway, this adoption strongly depends on the compromise and capabilities of the teachers, as main active actors in this new academic approach. Given this,
asking and analyzing user's opinions with regard to this scenario will help reaping the expected benefits of the technology.

@@

Given the potential benefits that chatbots could bring to the classroom, the
focus of the % EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
unnamed European project, the main sponsor of this study, is to explore best practices and innovative use of chatbots,
and to create a learning community of educators in higher education institutions.
Researchers and teachers participating in this project have developed several studies. Most of these works have been focused on how the
messaging applications and/or chatbots are normally applied to deliver personalized learning in classrooms that occurs anytime anywhere,
promote collaborative learning experiences, group discussions \cite{panah2020study} and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}.
The majority of these studies were conducted focusing on the teachers' perspectives, so this paper aims to provide more conclusions regarding
their opinions and needs, as well as the existing challenges or opportunities that arise when adopting these technologies in classrooms. This will lead to clarify
how they can positively contribute to enhancing the learning process in higher education institutions.

In this study, data collection involved two phases:
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their
  students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who they like to be
  with in the class messaging groups, and their expectations of chatbots
  in assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  familiar applications, such as Telegram or WhatsApp.
  Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed\textcolor{black}{; this will be detailed in Section \ref{sec:meth}}. The responses' analysis
led to the preparation of a second survey for teachers and 200 responses were
collected\textcolor{black}{, again following the methodology that will be explained later. The questions of the second survey focused on asking about teachers'
needs, opinions and preferences of the development of future technology-enhanced
tools and its potential impact on educational institutions policies. In order to assess these, this paper layed out the following research questions}:

@@

We will try to answer these research questions through the analysis of the two aforementioned surveys.
Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.

@@

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries, the most popular MIM platform
used by educators to provide assignments' feedback to students, support course
discussions, and contribute resources in informal learning settings
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

Several studies have been conducted to evaluate chatbots in higher education. For
instance, a thorough study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, with a focus on identifying characteristics and quality metrics such as
language, subject matter, and platform. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which responds to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada had
Elvira, a built-in chatbot in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the synchronization rather than making the updated 
of answered questions simpler, more interactive, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

Contrarily to the mentioned chatbots that are focused on administrative tasks, teaching-oriented chatbots are more sophisticated, as they
set personalized learning outcomes and monitor learning progress. For instance,
\cite{fernoagua2018intelligent} reported on ``eduAssistant'', a virtual teaching
assistant chatbot developed on the Telegram messaging platform. In this study,
the Telegram platform was chosen because it is easy to use, students are familiar
with its features, and it enables them to exchange messages in different formats
(text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram
can operate on all devices and operating systems. The ``eduAssistant'' chatbot
acts as an automatic agent in the teacher-content-student relationship, providing real-time
feedback loops and a personalized learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose questions to
their students and the chatbot assists those who need further help by giving
them more hints and reporting it to the educator's dashboard
\cite{fernoagua2018intelligent}. This can help educators identify
students who need more attention and provide them with more educational resources
relevant to their academic achievement. %In addition, a recent study
%has analysed how the use of chatbots positively affected the learning outcomes
%of students in a Chinese class \cite{doi:10.1177/0735633120929622}.

Despite its (arguably) successful implementation in different higher education
institutions, particularly on the fringes of educational activity rather than actual
student-teacher interaction, its implementation or deployment is not trivial.
Some authors, for instance \cite{sjostrom2018designing}, have proposed a
conceptual architecture for the adoption of teaching-oriented chatbots in higher
education.  This conceptual architecture is based on a systematic literature
review of previous studies examining the design of chatbots in higher
education, as well as making a content analysis of student emails and discussion
forum posts from four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasized the importance
of developing chatbots in platforms that students and educators are familiar
with and can easily access (i.e. Facebook Messenger), which was confirmed by
\cite{hobert2019you} and \cite{fernoagua2018intelligent}. In addition, \cite{sjostrom2018designing} argued that 
conceptualizing learners' questions could help designers integrate the
appropriate types of questions that chatbots should support for different
courses. Other authors, such as \cite{coronado2018cognitive}, have proposed agents that
store learning materials to be provided on demand to students, while \cite{crockett2017predicting} 
reported on tutoring systems that can
perform initial assessments of students' understanding and provide learning
materials to advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many
studies have focused on the evaluation of technology acceptance and usability
\cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain
where, according to \cite{hobert2019you}, specific pedagogical factors
such as learning success and increased motivation are more important.  Therefore,
to develop effective chatbots for higher education, the needs of all stakeholders
(i.e. educators, students, institutions, etc) should be carefully
collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}. These needs include, but are not limited to, student learning success due to higher motivation, but these are a posteriori effects that cannot be assessed in advance. This is why we focus in this paper on what  teachers are looking for in a Technology Acceptance Model (TAM) and associated technologies like chatbots. Both authors focus on what teachers need in terms of the latter; this paper will focus on a wider perspective, trying to determine what they are looking for in terms of general messaging technology. The classical literature \cite{tam91} already proposes that any adoption of technology must be tuned to the user needs and experience. In this paper, we will try to find out what those are in order to propose a model to facilitate the adoption of chatbot technology.

In line with this research, our previous work \cite{MoraChatbots2021:anon}
aimed at analyzing the expectations of students in this regard. Recent papers
have also analysed how the use of chatbots affected the learning outcomes of
students in a Chinese class \cite{doi:10.1177/0735633120929622}. This paper
analysed conversational chatbots in a one-to-one setting, finding that learning
truly benefited from it. This is, once again, an evidence of the benefit of
chatbots in certain settings; however, there are some prior experiences, as well as needs, that may
prevent the successful acceptance of the technology, and these are what we are
trying to find out in this paper, alng with what kind of features would improve
its acceptance.

@@

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Kind of interactions} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
%\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 171 & 60.6 & 110 & 39.0 \\
%\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers of the School/Faculty\end{tabular} & 13 & 4.6 & 269 & 95.4 \\
%\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 51 & 18.1 & 231 & 81.9 \\
%\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 113 & 40.1 & 169 & 59.9 \\ \hline
\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 166 & 59.1 & 115& 40.9 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers of the School/Faculty\end{tabular} & 19 & 6.7 & 266 & 94.3 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 58 & 20.5 & 224 & 79.4 \\
\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 81 & 28.7 & 201 & 71.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Kind of interactions preferred}
\label{tab:kindinteractions}
\end{table}

@@

%A total of 205 teachers responded to the second survey, 124 males (60.5\%), 65 females (31.7\%) and 16 individuals who chose not to identify their gender (7.8\%). The majority of the teachers were aged 45-55 years old (70; 34.1\%), followed by individuals aged 35-45 (67, 32.7\%), and 25-35 (42; 20.5\%), while the minority was aged >55 years old (26; 12.7\%). In terms of teaching experience, 59 teachers (28.8\%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9\%), 0-5 years (51; 24.9\%) and more than 25 years of experience (44; 21.5\%). The teachers were asked about what kind of interaction media features they value the most; the frequencies and percentages of their responses are given in the table that follows.

@@

%A total of 205 teachers responded to the second survey, 124 males (60.5\%), 65 females (31.7\%) and 16 individuals who chose not to identify their gender (7.8\%). The majority of the teachers were aged 45-55 years old (70; 34.1\%), followed by individuals aged 35-45 (67, 32.7\%), and 25-35 (42; 20.5\%), while the minority was aged $>$~55 years old (26; 12.7\%). In terms of teaching experience, 59 teachers (28.8\%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9\%), 0-5 years (51; 24.9\%) and more than 25 years of experience (44; 21.5\%). The teachers were asked about what kind of interaction media features they value the most \textcolor{black}{; i.e. what their preferred features are in the applications/tools that they use to interact with their students.} The frequencies and percentages of their responses are given in Table \ref{tab:interaction}.

@@

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers;
apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the mind of the
surveyed.  \textcolor{black}{In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app based messaging
systems vs. web-based messaging systems integrated in another learning management system).}

This inclusion of university-provided messaging applications in
the survey might explain why there are so few, just 6.7\%, that answer they do not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
people might not use it, but use some extraofficial or opt-in channels such as the other included in the survey; at the end, between these official and extraofficial channels most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant difference. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions that can arise from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots and just a few subjects selected
the option "Others". This may suggest  that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive, and if the responders were more acquainted with the possibilities of the technology, possibly this response would have shown up many more times. Another way to look about it is
that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burn out to think about
innovation.

We can speculate that their opinions towards chatbots are informed by negative
experiences with customer support chatbots, which is what most people will have
experienced. This will probably require further study, but also will need to be
taken into account when introducing chatbot technologies in the classroom:
it will have to be as simple, and as satisfactory, as possible,
to (possibly) change this existing perception.
